---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# Development Mode

mase is still under development.  Please use at your own risk!  

# mase

mase contains a collection of model-assisted generalized regression estimators (the post-stratification estimator, the ratio estimator, the linear and logistic regression estimator, the elastic net regression estimator, and the regression tree estimator) for finite population estimation of a total or mean from a single stage complex sample design.  It also contains the Horvitz-Thompson estimator and several variance estimators.

## Installation

You can install mase from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("Swarthmore-Statistics/mase")
```

## Example

Here's an example fitting the Horvitz-Thompson estimator:

```{r example}

library(mase)

## Estimates the mean and total of the api00 variable using the apisrs dataset in the survey package
library(survey)
data(api)
horvitzThompson(y = apisrs$api00, pi = apisrs$pw^(-1), var_est = TRUE, var_method = "HTSRS")
```

Please check out the vignette for more examples.

## Current Known Issues

* greg breaking down when have a single auxiliary variable. (Kevin/Shirley)

* Expanding the variance estimation options. (Mickey)

* Regression tree estimator in development. 

* Handling of categorical variable for y for all functions.  (Shirley)

* Ratio estimator variance broken? 

* Restructuring arguments of functions