---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# Development Mode

`mase` is still under development.  Please use at your own risk!  

# mase

`mase` contains a collection of model-assisted generalized regression estimators (the post-stratification estimator, the ratio estimator, the linear and logistic regression estimator, the modified linear and logistic regression estimator, the elastic net regression estimator, and the regression tree estimator) for finite population estimation of a total or mean from a single stage, unequal probability without replacement design.  It also contains the Horvitz-Thompson estimator and several variance estimators.

## Installation

You can install `mase` from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("mcconvil/mase")
```

## Example

Here's an example of fitting the Horvitz-Thompson estimator using Forestry data in Idaho. The data comes from the Forestry Inventory & Analysis department (FIA). 

```{r example1, message=F}
library(mase)
library(dplyr)

data(IdahoSamp)
data(IdahoPop)

samp <- filter(IdahoSamp, COUNTYFIPS == 16055) 
pop <- filter(IdahoPop, COUNTYFIPS == 16055) 

horvitzThompson(y = samp$BA_TPA_ADJ,
                N = pop$npixels,
                var_est = TRUE,
                var_method = "LinHTSRS")
```


We can also fit a linear regression estimator using that same data:

```{r example2, message = F}
xsample <- select(samp, tcc, elev) %>%
  as.data.frame()

xpop <- select(pop, names(xsample))

greg_est <- greg(y = samp$BA_TPA_ADJ,
     N = pop$npixels,
     xsample = xsample,
     xpop = xpop,
     var_est = TRUE,
     var_method = "LinHB",
     datatype = "means")
```

We still get the population total and mean estimates along with their variance estimates:

```{r}
greg_est[1:4]
```

But with this estimator we also get the weights

```{r}
greg_est[5]
```

and the coefficients for the model

```{r}
greg_est[6]
```

