---
title: "Mase Vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{mase_vin}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
library(tidyverse)
#devtools::install_github("mcconvil/mase")
library(mase)
knitr::opts_chunk$set(
  fig.width = 5,
  fig.align = "center",
  cache = TRUE,
  collapse = TRUE,
  comment = "#>"
)
```
The R package `mase` contains a set of model assisted survey estimation functions for estimating population estimates, subject to a complex sampling design. In particular, `mase` offers a multitude of model assisted Generalized Regression Estimators (GREGs), as well as  other direct estimators such as the ratio estimator for variances and the post-stratified and Horvitz-Thompson estimators for population totals.

To install `mase`, simply run the `R` code: `devtools::install_github("mcconvil/mase")`, and then attach it to your library. `tidyverse` is not required by `mase`, but will be used for the purpose of this vignette:

```{r install, eval = FALSE}
library(mase)
library(tidyverse)
```

The primary dataset we will be using for this vignette is a spatially indexed satellite raster dataset of forest attributes in the Black Hills in South Dakota from the U.S. Forest Service's Geodata Clearinghouse (https://data.fs.usda.gov/geodata/rastergateway/). The sample dataset is a subset of these spatially indexed raster data, and they represent a survey sample.

```{r load data}
data("black_hills_pop", "black_hills_samp", package = "mase")
head(black_hills_pop)
```
The population dataset contains the variables `forest_group`, `forest_nonforest_probability` and spatial grid indicies `s1` and `s2`, which represent longitudinal and latitudinal dimensions respectively. The `forest_group` variable 
```{r}
head(black_hills_samp)
```


```{r}
#prepare data
biomass <- black_hills_samp %>%
  pull(biomass)
x_samp <- black_hills_samp %>%
  dplyr::select(forest_group, forest_nonforest_probability, s1, s2)
# inclusion probs
inclusion_probs <- rep(c(0.0055, 0.00325, 0.00128128, 0.00721010), 58)
```

```{r}
m_enet <- gregElasticNet(y = biomass, x_sample = x_samp, x_pop = black_hills_pop,
                       pi = inclusion_probs, standardize = TRUE, var_est = TRUE)
m_enet$coefficients
plot(m_enet, y_sample = biomass, x_sample = x_samp, x_pop = black_hills_pop,
     spatial_1 = "s1", spatial_2 = "s2") +
  labs(x = "", y = "", fill = "Est. Biomass")
```

```{r random forest}
m_forest <- gregForest(y = biomass, x_sample = x_samp, x_pop = black_hills_pop,
                       pi = inclusion_probs, ntrees = 300, mtry = 2, var_est = TRUE)
m_forest
plot(m_forest, y_sample = biomass, x_sample = x_samp, x_pop = black_hills_pop,
     spatial_1 = "s1", spatial_2 = "s2") +
  labs(x = "", y = "", fill = "Est. Biomass")
predict(m_forest, x_samp)
```


```{r RBF}
m_rbf <- gregRBF(y = biomass, x_sample = x_samp, x_pop = black_hills_pop,
                         pi = inclusion_probs, lambda = 2, standardize = T, var_est = T)
m_rbf
plot(m_rbf, y_sample = biomass, x_sample = x_samp, x_pop = black_hills_pop,
     spatial_1 = "s1", spatial_2 = "s2") +
  labs(x = "", y = "", fill = "Est. Biomass")
predict(m_rbf, x_samp)
```


```{r neural nets}
m_nn <- gregNeuralNet(y = biomass, x_sample = x_samp, x_pop = black_hills_pop,
                      pi = inclusion_probs, standardize = T, lr = .005, epochs = 30,
                      H = 128, verbose = 0, var_est = TRUE)
m_nn
plot(m_nn, y_sample = biomass, x_sample = x_samp, x_pop = black_hills_pop,
     spatial_1 = "s1", spatial_2 = "s2")  +
  labs(x = "", y = "", fill = "Est. Biomass")

```





